---
title: "Mosquito-mediated epidemiological effects of malarone treatment"
number-sections: true
format:
  html:
    toc: true
    toc-depth: 4
editor: source
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: the-american-naturalist.csl
---

```{r include = FALSE, eval = FALSE}
knitr::purl("malarone.qmd", "tmp.R", documentation = FALSE)
source("tmp.R")
file.remove("tmp.R")
```

```{r include = FALSE, purl = FALSE}
par2 <- function(...) par(..., mgp = c(1.5, .5, 0), bty = "n")

knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par2(plt = c(.105, .97, .15, .95)) else NULL
  },
  margin2 = function(before, options, envir) {
    if (before) par2(plt = c(.22, .97, .15, .95)) else NULL
  },
  margin3 = function(before, options, envir) {
    if (before) par2(plt = c(.25, .95, .19, .9)) else NULL
  },
  margin3b = function(before, options, envir) {
    if (before) par2(plt = c(.25, .96, .17, .9)) else NULL
  },
  margin3c = function(before, options, envir) {
    if (before) par2(plt = c(.18, .89, .17, .9)) else NULL
  },
  margin4 = function(before, options, envir) {
    if (before) par2(plt = c(.05, .92, .15, .95)) else NULL
  },
  margin5 = function(before, options, envir) {
    if (before) par2(plt = c(.15, .95, .15, .95)) else NULL
  })

eps <- .8
knitr::opts_chunk$set(margin1    = TRUE,
                      fig.retina = 2,
                      fig.align  = "center",
                      fig.height = eps * 5, # default is 5
                      fig.width = eps * 7) # default is 7
```

In this analysis we use the formalism of @smith_statics_2004.

## Parameters

The path to the data:

```{r}
path2data <- paste0(Sys.getenv("HOME"), "/Library/CloudStorage/",
                    "OneDrive-OxfordUniversityClinicalResearchUnit/",
                    "GitHub/choisy/malarone/")
```

```{r include = FALSE}
make_path <- function(x) paste0(path2data, "cache/", x)
file_exists <- function(x) file.exists(make_path(x))
readRDS3 <- function(x) readRDS(make_path(x))
saveRDS2 <- function(object, file) saveRDS(object, make_path(file))
```

Whether or not to force the effects to be equal to $0$ at $t = 28$ days (in practice it
does not change things much anyway):

```{r}
t28_0 <- TRUE
```

Colors:

```{r}
YlOrRd10 <- hcl.colors(10, "YlOrRd", rev = TRUE)
YlOrRd11 <- hcl.colors(11, "YlOrRd", rev = TRUE)
color1 <- 2
color2 <- 3
color3 <- 4
```

Line width:

```{r}
lwd_val <- 2
```

Plot labels:

```{r}
xlab_pda <- "time post drug administration (hours)"
ylab_delta <- bquote(R[0] ~ "reduction factor" ~ delta)
ylab_pi <- bquote(R[0] ~ "reduction factor" ~ pi)
ylab_rho <- bquote(R[0] ~ "reduction factor" ~ rho)
ylab_b <- bquote("reducing factor on b and" ~ R[0])
```

Plots parameters sets:

```{r}
mcx <- 3 / 2
parmfrow3x3 <- function() {
  par(mfrow = c(3, 3), mex = mcx, cex.axis = mcx, cex.lab = mcx, cex.main = mcx)
}
```


## Packages

Required packages:

```{r}
required_packages <- c("readr", "readxl", "dplyr", "purrr", "stringr", "tibble")
```

Installing those that are not installed:

```{r}
to_inst <- required_packages[! required_packages %in% installed.packages()[,"Package"]]
if (length(to_inst)) install.packages(to_inst)
```

Loading some for interactive use:

```{r message = FALSE}
library(dplyr)
library(purrr)
```


## Functions

Tuning some base functions:

```{r}
seq2 <- function(...) seq(..., le = 512)
abline2 <- function(...) abline(..., col = "grey")
lines2 <- function(...) lines(..., lwd = lwd_val)
lines3 <- function(...) lines2(..., col = color1)
segments2 <- function(...) segments(..., col = color3, lwd = lwd_val)
plot2 <- function(...) plot(..., col = color3, pch = 3, lwd = lwd_val)
plotl <- function(...) plot(..., type = "l", col = color1, lwd = lwd_val)
arrows2 <- function(...) arrows(..., col = color2, length = .15, lwd = lwd_val)
readRDS2 <- function(f, ...) tibble::as_tibble(readRDS(paste0(path2data, f), ...))
legend2 <- function(...) legend(bty = "n", lty = 1, lwd = lwd_val, ...)
```

A function that replaces the value at index `i` in a vector `v` by the minimum of the
values of `v` with an index equal or higher than `i`:

```{r}
min_of_following <- function(x) {
  c(min(x), map_dbl(1:(length(x) - 1), ~ min(x[-(1:(.x))])))
}
```

A function that replace the `NA` values of a vector by `0`:

```{r}
NAto0 <- function(x) {
  x[is.na(x)] <- 0
  x
}
```


## Theoretical results

In this section we pull out the most relevant theoretical results from
@smith_statics_2004.

The **individual vectorial capacity** (noted $IC$ in @smith_statics_2004 but noted $C$
here) is defined by @saul_cyclical_1990 as the expected number of infectious bites from
a single vector after feeding on an infectious host and reads:

$$
C = c \cdot P_e \cdot S
$$

where $c$ is the **transmission efficiency from human to mosquito** (*i.e.* the
probability of becoming infected after feeding on an infectious human), $P_e$ is the
**probability of becoming infectious for an infected mosquito**:

$$
P_e = e^{-gn}
$$
(with $g$ the **mosquito death rate** and $n$ the **EIP**), and $S$ is the
**stability index**, *i.e.* the total number of bites per mosquito during its lifetime:

$$
S = \frac{a}{g}
$$

with $a$ the **mosquito biting rate**, *i.e.* the number of bites per mosquito.
Gathering everything gives:

$$
C = c \cdot e^{-gn}\frac{a}{g}
$$

The **basic reproduction ratio** is the expected number of infected humans per infected
human or, equivalently, the expected number of infected mosquitoes per infected 
mosquito and can be expressed as:

$$
R_0 = \frac{C \cdot b \cdot a \cdot m}{r}
$$ {#eq-R0}

where $C$ is the above-defined individual vectorial capacity, $b$ is the
**transmission efficiency from mosquito to human**, $r$ is the
**recovery rate of humans**, and $m$ is the **ratio of mosquitoes to humans**.

Finally, the **prevalance** $\bar{X}$ (*i.e.* proportion of infected humans) at
equilibrium can be expressed as a function of $R_0$:

$$
\bar{X} = \frac{R_0 - 1}{R_0 + c S}
$$

Note the main assumption of this framework:

**Assumption 1:** the mosquito death rate is independent of the mosquito’s age.


## Experimental data

The experimental results suggest that malarone

1. decreases by the multiplicative factor $1 - \alpha$ ($0 \leq \alpha \leq 1$) the
number of mosquitoes with oocysts at 7dpi (with $\alpha$ shown on the $y$ axis of
Figure 2a);
2. decreases by the multiplicative factor $1 - \beta$ ($0 \leq \beta \leq 1$) the
oocysts load in mosquitoes still harboring oocysts at 7dpi (with $\beta$ shown on
the $y$ axis of Figure 2b);
3. decreases by the multiplicative factor $1 - \gamma$ ($0 \leq \gamma \leq 1$) the
size of oocysts at 7dpi (with $\gamma$ shown on the $y$ axis of Figure 3a);
4. has no significant effect on the mosquito biting rate $a$ (Supplementary Figure 10);
5. has no significant effect on the mosquito death rate $g$ (Supplementary Figure 10).

Below we insert the 3 significant effects $\alpha$, $\beta$ and $\gamma$ into the
theoretical equations of the previous section.


### Mosquitoes with oocysts

Below, we generate the function $\alpha(t)$ from the experimental data of Figure 2a.
First we need a function that reads experimental data for the treatment group:

```{r}
make_reading_exp_data <- function() {
  hours <- c(0, 5, 48, 72)
  days <- c(5, 7, 14, 28)
  hash <- c(setNames(hours, paste0("H", hours)),
            setNames(24 * days, paste0("D", days)))
  function(file) {
    file |>
      readRDS2() |> 
      filter(treatment == "AL+AP") |> 
      mutate(time_hours = hash[stringr::str_remove(time_points, "0")],
             across(Inhibition, ~ .x / 100))
  }
}

reading_exp_data <- make_reading_exp_data()
```

We also need a function that converts such experimental data into a function:

```{r}
data2function <- function(data, correction = t28_0) {
  apply_correction <- function(x) {
    if (correction) return(mutate(x, median_val = c(head(median_val, -1), 0)))
    x
  }
  
  data |> 
    group_by(time_hours) |> 
    summarise(median_val = max(0, median(Inhibition))) |> 
    bind_rows(tibble(time_hours = 0, median_val = 0)) |> 
    arrange(time_hours) |> 
    apply_correction() |> 
    with(approxfun(time_hours, median_val))
}
```

where `correction = TRUE` means that the value at $t = 28$ days is forced to be equal
to $0$. Let's now pull out the data from Figure 2a:

```{r}
#alpha_data <- reading_exp_data("percentage_inhibition9.rds")
alpha_data <- reading_exp_data("figure2a.rds")
```

and create a function `alpha()` of the percentage $\alpha(t)$ inhibition on sporozoite
infection rate as a function of time $t$ post drug administration:

```{r}
alpha0 <- data2function(alpha_data)
alpha <- function(x) NAto0(alpha0(x))
```

Let's now visualize this function. For that we first need a function that plots
experimental data:

```{r}
plot_experimental_data <- function(data, ...) {
  with(data, plot2(jitter(time_hours), Inhibition, xlab = xlab_pda, ...))
  abline(h = 0)
}
```

With this function `plot_experimental_data()` defined, this is what the $\alpha(t)$
function coded in `alpha()` looks like together with the experimental data:

```{r}
plot_experimental_data(alpha_data, ylim = c(-1, 1),
                       ylab = bquote("reduction" ~ alpha ~
                                       "of mosquitoes with oocysts"))
pda_t <- seq2(0, 700)
lines3(pda_t, alpha(pda_t))
```

The reduction factor $1 - \alpha$ of the number of mosquitoes with oocysts (Figure
2a) converts directly into a transmission-blocking activity (TBA) of similar magnitude
on the mean individual vectorial capacity:

$$
(1 - \alpha(t))\cdot C
$$

where $C$ is the vectorial capacity in absence of malarone and $t$ is the time post
drug administration. Similarly, @eq-R0 shows that the reduction of $R_0$ is

$$
(1 - \alpha(t))\cdot R_0
$$

where $R_0$ is the basic reproduction ratio in absence of malarone.

### Number of oocysts

Let's pull out the data from Figure 2c:

```{r}
beta_data <- reading_exp_data("figure2c.rds")
```

and create a function `beta(t)` on the percentage $\beta(t)$ inhibition on oocysts
load as a function of time $t$ post drug administration:

```{r}
beta0 <- data2function(beta_data)
beta <- function(x) NAto0(beta0(x))
```

And let's visualize this function $\beta(t)$ together with the experimental data:

```{r}
#| label: fig-beta_data
#| fig-cap: experimental data and model from Figure 4B.

plot_experimental_data(beta_data, ylim = c(-1, 1),
                       ylab = bquote("reduction" ~ beta ~
                                       "of the number L of oocysts"))
betas <- beta(pda_t)
lines3(pda_t, betas)
```

The insertion of the $\beta$ effect into @eq-R0 is less direct and straightforward
than for $\alpha$. To do so, we need to understand how a reduction in the oocysts
load $L$ translates into a reduction of $b$ in @eq-R0.

Figure 2B from @aleshnick_experimental_2020 suggests a non-linear relationship $h$
between the sporozoite load and the transmission efficiency $b$ of *P. yoelii* from
*An. stephensi* mosquito to mice with 10,000 salivary gland sporozoites being
significantly more likely to initiate a malaria infection. Furthermore, Figure 1 from
@miura_strong_2019 documents a strong linear relationship between the numbers of
*P. falciparum* oocysts at 8 dpi and sporozoites at 16 dpi in *An. stephensi*
mosquitoes with about 1,000 sporozoites per oocyst (966, 95\% CI: 893–1038)). The
effect $\rho$ of $\beta$ on $b$ can thus be modelled as so:

$$
\xi(\beta, L) = \frac{h((1 - \beta)L)}{h(L)}
$$ {#eq-TRA}

where $h$ is the functional relationship between the oocsyts load $L$ and the
mosquito-to-human transmission efficiency $b$ derived from Figures 2B of
@aleshnick_experimental_2020 and 1 of @miura_strong_2019. We can compute a population
mean value of this reduction using the distribution of the oocyst load in the
population as so

$$
f(\beta) = \sum_{L = 1}^{+\infty}\xi(\beta, L) \omega(L)
$$ {#eq-TRA_mean}

where $\omega(L)$ is the frequency of load $L$ in the population. Let's thus first look
at the distribution of the oocyst loads $L$ from the control data:

```{r}
oocysts_control <- path2data |>
  paste0("ASAAP4b_oocyst2.txt") |>
  readr::read_table() |> 
  filter(day_of_dissection == "7dpi",
         time_points %in% c("Control", "Unfed"),
         total_oocysts > 0) |> 
  dplyr::select(total_oocysts) |> 
  na.exclude()

total_oocysts <- oocysts_control$total_oocysts
  
hist(total_oocysts, main = NA,
     xlab = "oocyst load L in the controls", ylab = "frequency")
```

An alternative histogram:

```{r}
max_load <- max(total_oocysts)

hist(total_oocysts, 0:max_load + .5, main = NA,
     xlab = "oocyst load L in the controls", ylab = "frequency")
```

The cumulative distribution together with a fitted geometric distribution:

```{r}
oocysts_control |> 
  group_by(total_oocysts) |> 
  tally() |> 
  mutate(cum_n = cumsum(n) / sum(n)) |> 
  with(plot(total_oocysts, cum_n, type = "s", ylim = 0:1,
            xlab = "oocyst load L", ylab = "cumulative proportion of samples"))

p_par <- 1 / mean(total_oocysts)
x <- 1:max_load
lines(x, cumsum(dgeom(x, p_par)), type = "s", col = 2)
```

Now that we have a calibrated parametric distribution of the oocyst loads $L$, let's
look at the data from @aleshnick_experimental_2020 converted to oocyst load using the
1,000 factor of @miura_strong_2019:

```{r}
library(mgcv)

miura_factor <- 1000

aleshnick2020 <- path2data |> 
  paste0("ppat.1008181.s006.xlsx") |> 
  readxl::read_excel() |> 
  rename(infection = `Malaria Infection`, load = `Salivary Gland Load`) |> 
  dplyr::select(load, infection)

mod <- gam(infection ~ s(load, k = 50), binomial, aleshnick2020, method = "REML")
x <- seq2(1, max(oocysts_control$total_oocysts) * miura_factor)
x_ooc <- x / miura_factor
predictions <- predict(mod, list(load = x), se.fit = .95, type = "response")

plot(x_ooc, predictions$fit, ylim = c(0, max(predictions$ul)), , type = "n",
     xlab = "oocyst load L", ylab = "infection probability f")
polygon(c(x_ooc, rev(x_ooc)), c(predictions$ul, rev(predictions$ll)),
        border = NA, col = adjustcolor(4, .2))
lines(x_ooc, predictions$fit, lwd = 3, col = 4)

aleshnick2020 |> 
  arrange(load) |> 
  mutate(across(load, ~ .x / 1000),
         group = c(rep(1:10, each = 41), 10, 10)) |> 
  group_by(group) |> 
  summarise(min_load   = min(load),
            max_load   = max(load),
            mean_load  = mean(load),
            proportion = list(binom.test(sum(infection), length(infection)))) |> 
  mutate(estimate = map_dbl(proportion, "estimate"),
         confint = proportion |>
                     map("conf.int") |>
                     map(~ setNames(.x, c("ll", "ul")))) |> 
  tidyr::unnest_wider(confint) |> 
  with({
    arrows(mean_load, ll, mean_load, ul, .1, 90, 3)
    points(mean_load, estimate, pch = 19)
    abline(v = min_load, col = "grey")
  })

aleshnick2020 |> 
  mutate(across(infection, ~ .x * .9)) |> 
  with(points(load / miura_factor, infection, pch = "|"))
```

Let's make a simplified version of it:

```{r}
x_threshold <- 25000
x_threshold2 <- x_threshold / miura_factor

plateau_val <- aleshnick2020 |> 
  filter(load > x_threshold) |> 
  with(sum(infection) / length(infection))

slope_val <- miura_factor * plateau_val / x_threshold

h <- function(x) {
  map_dbl(x, function(x) {
    if (x < x_threshold2) return(slope_val * x)
    plateau_val
  })
}
```

And this is what this function $h$ function coded in `h()` looks like:

```{r}
L_val <- seq2(0, 100)
plotl(L_val, h(L_val),
      xlab = "oocysts load L", ylab = "infection probability h")
```

The same on the whole range of oocyst load:

```{r}
L_val <- seq2(0, max_load)
plotl(L_val, h(L_val),
      xlab = "oocysts load L", ylab = "infection probability h")
```

Combining `h()` with the observed distribution of oocyst loads in order to convert a
value of $\beta$ into a reduction factor of $b$ or $R_0$:

```{r}
f <- function(beta, prob = p_par, loads = 1:max_load) {
  densities <- dgeom(loads, prob)
  densities <- densities / sum(densities)
  denominator <- h(loads)
  map_dbl(beta, ~ sum(densities * h((1 - .x) * loads) / denominator))
}
```

And here is the TRA as a function of $\beta$:

```{r fig.height = eps * 6, fig.width = eps * 8, margin1 = FALSE, margin4 = TRUE}
#| label: fig-b_reduction
#| fig-cap: the conversion of a reduction on the sporozoite load $L$ into a reduction
#| of the mosquito-to-human probability of transmission $b$ and $R_0$.

beta_vals <- seq2(0, 1)
y_vals <- f(beta_vals)
opar <- par(pty = "s", xpd = TRUE)
plot(1 - beta_vals, y_vals, type = "l", xaxs = "i", yaxs = "i", ylab = ylab_b,
     xlab = bquote("reducing factor 1 -" ~ beta ~ "on oocyst load L"))
par(opar)
```

By combining @fig-beta_data and @fig-b_reduction we obtain the reduction factor of $b$
and $R_0$ as a function of time $t$ post drug administration:

```{r}
plot(NA, xlim = c(0, 700), ylim = 0:1, xlab = xlab_pda, ylab = ylab_b)
lines(pda_t, f(betas))
```


### Sizes of oocysts at 7dpi

Malarone decreases by the multiplicative factor $1 - \gamma$ ($0 \leq \gamma \leq 1$)
the size of oocysts at 7 dpi. Furthermore, figure 5B from @werling_steroid_2019
suggests a linear relationship between the development time and the oocyst size. By
assuming that the EIP corresponds to a given size of the oocysts, we can model the
effect of $\gamma$ on $n$ as so:

$$
\frac{1}{1 - \gamma(t)}n
$$

With this established, let's now pull out the data from Figure 3a:

```{r}
gamma_data <- reading_exp_data("percentage_inhibition_size2.rds")
```

and create a function `gamma()` on the percentage $\gamma(t)$ decrease in oocysts
size at 7 dpi as a function of time $t$ post drug administration:

```{r}
gamma0 <- data2function(gamma_data)
gamma <- function(x) NAto0(gamma0(x))
```

And let's visualize this function $\gamma(t)$ together with the experimental data:

```{r}
plot_experimental_data(gamma_data, ylim = c(-.65, .65),
                       ylab = bquote("sporozoite size reduction factor" ~ gamma))
lines3(pda_t, gamma(pda_t))
```

**Assumption 2:** the EIP corresponds to a given size of the oocysts.


### Combining the 3 effects

In this section we combine the 3 effects reviewed above into a reduction effect of the
basic reproduction ratio $R_0$. If we assume that all the mosquitoes feed on people who
received malarone $t$ time units ago, the reduction factor $\delta(t)$ of $R_0$ reads:

$$
\delta(t, g, n, L) = (1 - \alpha(t))f(\beta(t))e^{-gn\gamma(t)/(1-\gamma(t))}
$$
where $L$ is the mean oocyst load in absence of malarone. The corresponding R code:

```{r}
delta <- function(t, g, n) {
  (1 - alpha(t)) * f(beta(t)) * exp(- g * n * gamma(t) / (1 - gamma(t)))
}
```

**Note 1:** $g$ and $n$ should be expressed in the same time units.  
**Note 2:** $t$ should be expressed in hours.

If we now assume that all the mosquitoes feed on people who were treated sometimes
during the past $D$ time units, then the mean reduction factor of $R_0$ reads:

$$
\lambda(g, n, L, D) = \frac{1}{D}\int_0^D\delta(t, g, n, L)dt
$$ {#eq-lambda}

Ideally, we'd like $D$ to be large enough so that $\delta(t = D, g, n, L)\simeq 0$. The
corresponding R code is:

```{r}
lambda <- function(D, g, n, ...) {
  integrate(function(x) delta(x, g, n), 0, D, ...)$value / D
}
```

**Note 1:** $g$ and $n$ should be expressed in the same time units.  
**Note 2:** $D$ should be expressed in hours.


## Epidemiological effects of malarone

In this section we use $\lambda$ to develop what interests us here: the
mosquito-mediated epidemiological effects of malarone. Since we are interested only in
the effect of malarone that is mediated by the mosquito compartment, in what follows
we'll disregard the direct and intended effect of malarone, which is the curation of 
infected people. We will also consider two types of use of malarone: mass treatment
where a given proportion of the population is treated independently of their clinical
status and treatment of the infected, which is the way that malarone is currently used.
We'll start however with the mass treatment as it's the easiest case to model and it
will provide us with a gold standard to which to compare the effects of the treatment
of infected.

### Mass treatment

Mass treatment is not affected by the epidemiological context so its mosquito-mediated
effect $\pi$ can be expresses as:

$$
\pi(\mu, g, n, L, D) = 1 - \left(1 - \min_{\Delta \ge D}\lambda\left(g, n, L, \Delta\right)\right)\mu
$$

where $\mu$ is the proportion of the population that receive the treatment and $1/D$ is
the frequency of treatment, assuming an absence of synchrony of treatment intake
between people (*i.e.* the times at which each individual takes its treatment are
random).

**Note:** $g$ and $n$ should be expressed in the same time units.

### Treatment of infected

Finally, let's put the effect of malarone into an epidemiological context. If $\bar{X}$
is the disease prevalence and $r$ the human recovery rate, then we can approximate the
disease incidence by $\bar{X}r$. If $\tau$ is the treatment coverage (*i.e.* the
proportion of new cases who receive treatment), then the treatment incidence is
$\tau\bar{X}r$. This means that the proportion of people who received treatment
sometimes during the past $D$ time units is $1 - e^{-\tau\bar{X}rD}$ and that the
mean reduction factor $\rho$ of $R_0$ reads:

$$
\rho(g, n, L, \tau, \bar{X}, r, D) = 1 - (1 - \lambda(g, n, L, D))\left(1 - e^{-\tau\bar{X}rD}\right)
$$ {#eq-rho}

**Note 1:** $g$ and $n$ should be expressed in the same time units.  
**Note 2:** $r$ and $D$ should be both expressed in hours.

## Numerical analysis

The table below lists the parameters of the models, together with their definitions and
their values and units:

| parameter | definition                        | values        | unit        |
|:----------|:----------------------------------|--------------:|------------:|
| $g$       | mosquito death rate               | 1 / 14        |  /day.      |
| $n$       | extrinsic incubation period (EIP) | 6, 10 or 14   | days        |
| $r$       | human recovery rate               | 1 / 14        | /day        |
| $1/D$     | frequency of treatment intake     | 1/28 - 1      | /day        |
| $\mu$     | mass treatment coverage           | 0-100         | %           |
| $\tau$    | infected treatment coverage       | 0-100         | %           |
| $\bar{X}$ | prevalence                        | 0-100         | %           |


### Reducing factor $\delta(t)$

The $R_0$ reducing factor $\rho(t)$ is computed by assuming that all the mosquitoes 
feed on people $t$ hours after drug administration. For a mosquito life expectancy
$1/g$ of 14 days and an EPI $n$ of 7 days, if the all the mosquitoes feed on people 100
hours after drug administration, the $R_0$ reducing factor $\delta$ is:

```{r}
(effect_at_t100 <- delta(t = 100, g = 1 / 14, n = 7))
```

More generally, if all the mosquitoes feed on people $t$ hours after drug 
administration, the $R_0$ reducing factor looks like:

```{r}
pda_t2 <- seq(0, 24 * 28, le = 2^10)
ys <- delta(t = pda_t2, g = 1 / (14 * 24), n = 7 * 24)
plotl(pda_t2, ys, ylim = 0:1, xlab = xlab_pda, ylab = ylab_delta)
abline(v = 100, col = 4)
abline(h = effect_at_t100, col = 4)
points(100, effect_at_t100, col = 4)
```

where the blue point and lines show the case where all the mosquitoes feed on people
100 hours after drug administration.


### Mass treatment

The $R_0$ reducing factor $\lambda(t)$ is computed by assuming that all the mosquitoes
feed on people that were treated sometimes during the past $D$ time units. Considering
that all the mosquitoes feed of people who received treatment in the past 28 days, the
$R_0$ reducing factor $\lambda$ is:

```{r}
(mean_effect_over28days <- lambda(D = 28 * 24, g = 1 / 14, n = 7))
```

which, on the previous plot, looks like:

```{r}
plot(pda_t2, ys, ylim = 0:1, xaxs = "i", yaxs = "i", type = "n",
     xlab = xlab_pda, ylab = ylab_delta)
polygon(c(pda_t2[1], pda_t2, tail(pda_t2, 1)), c(0, ys, 0),
        border = NA, col = adjustcolor(4, .2))
lines3(pda_t2, ys)
abline(h = mean_effect_over28days, col = 4, lty = 2, lwd = lwd_val)
text(mean(pda_t2), 1.04 * mean_effect_over28days, col = 4,
     bquote("        mean" ~ R[0] ~ "reduction factor" ~ lambda))
box(bty = "o")
```

A function that computes the values for $\lambda$:

```{r}
future::plan(multisession, workers = parallel::detectCores())

compute_lambda_val <- function(n, g, D) {
  tidyr::expand_grid(n, g, D) |> 
    mutate(lambda = furrr::future_pmap_dbl(list(D, g, n), ~ lambda(..1, ..2, ..3)))
}
```

Let's now compute $\lambda$ values (takes ca 25" on the laptop):

```{r eval = FALSE}
lambda_vals <- compute_lambda_val(n = c(6, 10, 14),
                                  g = 1 / c(14, 10, 7),
                                  D = c(seq(1, 7, .1), 8:28) * 24)
```

```{r include = FALSE}
if (file_exists("lambda_vals.rds")) {
  lambda_vals <- readRDS3("lambda_vals.rds")
} else {
  lambda_vals <- compute_lambda_val(n = c(6, 10, 14),
                                    g = 1 / c(14, 10, 7),
                                    D = c(seq(1, 7, .1), 8:28) * 24)
  saveRDS2(lambda_vals, "lambda_vals.rds")
}
```

A function that plots one panel:

```{r}
ats <- seq(30, 5, -5)
prev <- 0:10 / 10

plot_pi_D <- function(x, xlab = "intake frequency (/day)", ylab = ylab_pi) {
  plot(NA, xlim = c(30, 0), ylim = 0:1, axes = FALSE,
       xaxs = "i", yaxs = "i", xlab = xlab, ylab = ylab)
  abline(v = ats, col = "grey")
  abline(h = prev, col = "grey")
  box(bty = "o")
  walk2(prev, YlOrRd11,
        ~ with(x, lines(D / 24, (1 - (1 - min_of_following(lambda)) * .x),
                        col = .y, lwd = lwd_val)))
}

plot_pi_D2 <- function(...) plot_pi_D(..., xlab = NA)
plot_pi_D3 <- function(...) plot_pi_D(..., ylab = NA)
plot_pi_D4 <- function(...) plot_pi_D2(..., ylab = NA)
```

Some tunings:

```{r}
mtext2 <- function(x) mtext(paste("EPI =", x, "days"), 3, .5)
mtext3 <- function(x) mtext(paste0("g = 1/", x, " /day"), 4, .75)

legend4 <- function(...) {
  legend(lty = 1, lwd = lwd_val, adj = c(1, .5), x.intersp = 4, ...)
}

legend5 <- function(...) {
  par(xpd = TRUE)
  legend4(col = YlOrRd11, inset = c(-.25, -.23), ...)
  par(xpd = FALSE)
}

axis1 <- function() axis(1, c(ats, 1), c(paste0("1/", ats), 1))
```

The figure:

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
opar <- par(mfrow = c(3, 3))
plot_pi_D2(filter(lambda_vals, n ==  6, g <  .1))
mtext2(6); axis(2)
plot_pi_D4(filter(lambda_vals, n ==  6, g == .1))
mtext2(10)
legend5("bottomleft", title = "proportion treated", text.width = .5,
        legend = paste(seq(0, 100, 10), "%"))
plot_pi_D4(filter(lambda_vals, n ==  6, g >  .1))
mtext2(14); mtext3(14)
plot_pi_D2(filter(lambda_vals, n == 10, g <  .1))
axis(2)
plot_pi_D4(filter(lambda_vals, n == 10, g == .1))
plot_pi_D4(filter(lambda_vals, n == 10, g >  .1))
mtext3(10)
plot_pi_D(filter(lambda_vals, n == 14, g <  .1))
axis1(); axis(2)
plot_pi_D3(filter(lambda_vals, n == 14, g == .1))
axis1()
plot_pi_D3(filter(lambda_vals, n == 14, g >  .1))
axis1(); mtext3(7)
par(opar)
```


### Treatment of infected

```{r}
rho <- function(lambda, X, tau, r = 1 / (14 * 24)) {
  lambda |>
    filter(D > 27 * 24) |> 
    tidyr::expand_grid(X, tau) |> 
    mutate(rho = 1 - (1 - lambda) * (1 - exp(- tau * X * r * D)))
}
```

```{r}
rho_vals <- rho(lambda_vals, prev, seq2(0, 1))
```

A function that plots one panel:

```{r}
abline3 <- function(...) abline(..., col = "lightgrey")
abline4 <- function(...) abline3(..., lwd = lwd_val)

plot_tau_rho <- function(x, ylab = ylab_rho,
                         xlab = bquote(atop("Proportion" ~ tau ~ "of new cases",
                                            "that are treated (%)"))) {
  plot(NA, xlim = 0:1, ylim = 0:1, xlab = xlab, ylab = ylab, axes = FALSE)
  abline3(v = seq(0, 1, .05))
  abline4(v = seq(0, 1, .1))
  abline4(h = seq(0, 1, .1))
  abline3(h = seq(0, 1, .05))
  box(bty = "o")
  x |> 
    group_split(X) |> 
    walk2(YlOrRd11, ~ with(.x, lines(tau, rho, col = .y)))
}

plot_tau_rho2 <- function(...) plot_tau_rho(..., xlab = NA)
plot_tau_rho3 <- function(...) plot_tau_rho(..., ylab = NA)
plot_tau_rho4 <- function(...) plot_tau_rho2(..., ylab = NA)
```

The figure:

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
opar <- par(mfrow = c(3, 3))
plot_tau_rho2(filter(rho_vals, n ==  6, g <  .1))
mtext2(6); axis(2)
plot_tau_rho4(filter(rho_vals, n ==  6, g == .1))
mtext2(10)
legend5("bottomleft", title = "prevalence", text.width = 0, 
          legend = paste(seq(0, 100, 10), "%"))
plot_tau_rho4(filter(rho_vals, n ==  6, g >  .1))
mtext2(14); mtext3(14)
plot_tau_rho2(filter(rho_vals, n == 10, g <  .1))
axis(2)
plot_tau_rho4(filter(rho_vals, n == 10, g == .1))
plot_tau_rho4(filter(rho_vals, n == 10, g >  .1))
mtext3(10)
plot_tau_rho(filter(rho_vals, n == 14, g <  .1))
axis(1); axis(2)
plot_tau_rho3(filter(rho_vals, n == 14, g == .1))
axis(1)
plot_tau_rho3(filter(rho_vals, n == 14, g >  .1))
axis(1); mtext3(7)
par(opar)
```


### Examples with $R_0$

If we want to look at a specific example of $R_0$, we need to acknowledge that,
everything else being constant, $R_0$ is proportional to the probability $P_e$ of an
infected mosquito to become infectious:

$$
R_0 \propto e^{-gn}
$$

with this being considered, let's now consider for example a reference value of
$R_0 = 5$ in absence of treatment and when the EPI $n$ is equal to the mosquito life
expectancy $1/g = 14$ days. The effect of mass treatment would look like:

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_mass(lambda_vals, ylim = c(0, ymax),
                 ylab = bquote(R[0]), hlines = 0:ymax, R0o = 5)
```

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_mass(lambda_vals2, ylim = c(0, ymax),
                 ylab = bquote(R[0]), hlines = 0:ymax, R0o = 5)
```

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_mass(lambda_vals3, ylim = c(0, ymax),
                 ylab = bquote(R[0]), hlines = 0:ymax, R0o = 5)
```

And the effect of the treatment of the infected would look like:

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_infected(rho_val1, ylim = c(0, ymax), ylab = bquote(R[0]),
                     hlines1 = 0:ymax, hlines2 = seq(0, ymax, .5), R0o = 5)
```

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_infected(rho_val2, ylim = c(0, ymax), ylab = bquote(R[0]),
                     hlines1 = 0:ymax, hlines2 = seq(0, ymax, .5), R0o = 5)
```

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_infected(rho_val3, ylim = c(0, ymax), ylab = bquote(R[0]),
                     hlines1 = 0:ymax, hlines2 = seq(0, ymax, .5), R0o = 5)
```

### Main text figure

This is the figure that shows the mass treatment and the treatment of infected for
$L = 50,000$ and $n = 10$ days:

```{r fig.width = 8, margin1 = FALSE, margin5 = TRUE}
mtext4 <- function(...) mtext(..., line = -.5, cex = 2, font = 2)
at_val <- 4.7
xmax <- 30

leg <- paste(seq(0, 100, 10), "%")
nb_mid <- 3

opar <- par(mfrow = 1:2)

plot_mass_treatment(attr(lambda_vals, "params")$D_val / 24, lambda_vals[[5]])
legend4("bottomleft", title = "proportion treated", text.width = .5,
        col = c(YlOrRd11[1:3], rep("#FFFFFF", nb_mid), YlOrRd11[10:11]), 
        legend = c(leg[1:3], rep(".", nb_mid), leg[10:11]))
mtext4("a", at = at_val + xmax)

plot_treatment_infected(attr(rho_val1, "params")$tau, rho_val1[[5]])
legend4("bottomleft", title = "prevalence", col = YlOrRd11, text.width = 0,
        legend = leg)
mtext4("b", at = - at_val / xmax)

par(opar)
```


```{r fig.width = 8, margin1 = FALSE, margin5 = TRUE}
mtext4 <- function(...) mtext(..., line = -.5, cex = 2, font = 2)
at_val <- 4.7
xmax <- 30

leg <- paste(seq(0, 100, 10), "%")
nb_mid <- 3

opar <- par(mfrow = 1:2)

plot_treatment_infected(attr(rho_val1, "params")$tau, rho_val1[[5]])
legend4("bottomleft", title = expression(paste("prevalence ", bar(X))),
        col = YlOrRd11, text.width = 0, legend = leg)

mtext4("a", at = - at_val / xmax)

plot_mass_treatment(attr(lambda_vals, "params")$D_val / 24, lambda_vals[[5]])
legend4("bottomleft", title = "proportion treated", text.width = .5,
        col = c(YlOrRd11[1:3], rep("#FFFFFF", nb_mid), YlOrRd11[10:11]), 
        legend = c(leg[1:3], rep(".", nb_mid), leg[10:11]))

mtext4("b", at = at_val + xmax)

par(opar)
```

#### PDF output

Au format *Nature Medicine*:

```{r}
PDF <- TRUE

abline4pdf <- function(...) abline(..., col = "lightgrey")
abline3pdf <- function(...) abline4pdf(..., lwd = .5)

plot_treatment_infected_pdf <- function(x, y, x_axis = TRUE, y_axis = TRUE, ylim = 0:1,
                                        xlab = bquote("proportion" ~ tau ~ "of new cases that are treated"),
                                        ylab = ylab_rho,
                                        hlines1 = seq(0, 1, .1),
                                        hlines2 = seq(0, 1, .05),
                                        transform = 1) {
  plot(NA, xlim = 0:1, ylim = ylim, type = "n", xaxs = "i", yaxs = "i", axes = FALSE,
       ann = FALSE, xlab = xlab, ylab = ylab)
  ats <- seq(0, 1, .2)
  axis(1, ats, 100 * ats)
  axis(2, mgp = c(1, .75, 0))
  mtext(bquote("Proportion" ~ tau ~ "of new cases that are treated (%)"), 1, 1.5)
  mtext(ylab_rho, 2, 2, las = 0)
  abline3pdf(v = seq(0, 1, .05))
  abline4pdf(v = seq(0, 1, .1))
  abline4pdf(h = hlines1)
  abline3pdf(h = hlines2)
  walk2(y, YlOrRd11, ~ lines(x, .x * transform, col = .y))
  box(bty = "o")
}

legend4pdf <- function(...) {
  legend(lty = 1, lwd = 1, adj = c(1, .5), x.intersp = 4, ...)
}

plot_mass_treatment_pdf <- function(x, y, x_axis = TRUE, y_axis = TRUE, ylim = 0:1,
                                xlab = "intake frequency (every D days)",
                                ylab = ylab_pi,
                                hlines = seq(0, 1, .1), transform = 1) {
  plot(rev(x), y, xlim = c(xmax, 0), ylim = ylim, type = "n",
       xlab = xlab, ylab = ylab, xaxs = "i", yaxs = "i", axes = FALSE, ann = FALSE)
  axis(1)
  axis(2, mgp = c(1, .75, 0))
  mtext("Intake frequency (every D days)", 1, 1.5)
  mtext(ylab_pi, 2, 2, las = 0)
  abline(v = seq(0, xmax, 5), col = "grey")
  abline(h = hlines, col = "grey")
  walk2(seq(0, 1, .1), YlOrRd11,
        ~ lines(rev(x), rev((1 - (1 - min_of_following(y)) * .x) * transform),
                col = .y))
  box(bty = "o")
}

#######################################################################################

if (PDF) {
  pdf("figure6.pdf",
      width = 121 / 25.4, height = 60.5 / 25.4,
      family = "Helvetica", pointsize = 6)
}
  
mtext4pdf <- function(...) mtext(..., line = -.5, cex = 1.5, font = 2)
at_val <- 4.7
xmax <- 30

leg <- paste(seq(0, 100, 10), "%")
nb_mid <- 3

opar <- par(mfrow = 1:2, mgp = c(1.5, .5, 0), plt = c(.15, .95, .13, .93), las = 1)

plot_treatment_infected_pdf(attr(rho_val1, "params")$tau, rho_val1[[5]])
legend4pdf("bottomleft", title = expression(paste("Prevalence ", bar(X))), text.width = 0,
        col = YlOrRd11,
        legend = leg,
        bg = "white")

mtext4pdf("a", at = - at_val / xmax)

plot_mass_treatment_pdf(attr(lambda_vals, "params")$D_val / 24, lambda_vals[[5]])
legend4pdf("bottomleft", title = expression(paste("Proportion ", mu, " treated")), text.width = .5,
        col = c(YlOrRd11[1:5], rep("#FFFFFF", nb_mid), YlOrRd11[10:11]), 
        legend = c(leg[1:5], rep(".         .     ", nb_mid), leg[10:11]),
        bg = "white")

mtext4pdf("b", at = at_val + xmax)

par(opar)

if (PDF) dev.off()
```

### New frequencies

#### Supplementary figures

Let's now compute $\lambda$ values (takes ca 8'00"):

```{r eval = FALSE}
lambda_vals1b <- compute_lambda_val(D_val = (3 * 7):(8 * 7) * 24,
                                    n_val = c(6, 10, 14),
                                    L_val = c(1, 5, 10) * 1e4,
                                    g = 1 / 14)
```

(takes ca 8'00"):

```{r eval = FALSE}
lambda_vals2b <- compute_lambda_val(D_val = (3 * 7):(8 * 7) * 24,
                                    n_val = c(6, 10, 14),
                                    L_val = c(1, 5, 10) * 1e4,
                                    g = 1 / 10)
```

(takes ca 8'00"):

```{r eval = FALSE}
lambda_vals3b <- compute_lambda_val(D_val = (3 * 7):(8 * 7) * 24,
                                    n_val = c(6, 10, 14),
                                    L_val = c(1, 5, 10) * 1e4,
                                    g = 1 / 7)
```

```{r include = FALSE}
if (file_exists("lambda_vals1b.rds")) {
  lambda_vals1b <- readRDS3("lambda_vals1b.rds")
} else {
  lambda_vals1b <- compute_lambda_val(D_val = (3 * 7):(8 * 7) * 24,
                                      n_val = c(6, 10, 14),
                                      L_val = c(1, 5, 10) * 1e4,
                                      g = 1 / 14)
  saveRDS2(lambda_vals1b, "lambda_vals1b.rds")
}

if (file_exists("lambda_vals2b.rds")) {
  lambda_vals2b <- readRDS3("lambda_vals2b.rds")
} else {
  lambda_vals2b <- compute_lambda_val(D_val = (3 * 7):(8 * 7) * 24,
                                      n_val = c(6, 10, 14),
                                      L_val = c(1, 5, 10) * 1e4,
                                      g = 1 / 10)
  saveRDS2(lambda_vals2b, "lambda_vals2b.rds")
}

if (file_exists("lambda_vals3b.rds")) {
  lambda_vals3b <- readRDS3("lambda_vals3b.rds")
} else {
  lambda_vals3b <- compute_lambda_val(D_val = (3 * 7):(8 * 7) * 24,
                                      n_val = c(6, 10, 14),
                                      L_val = c(1, 5, 10) * 1e4,
                                      g = 1 / 7)
  saveRDS2(lambda_vals3b, "lambda_vals3b.rds")
}
```

```{r supfig_mass1, fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
multipanels_mass2(lambda_vals1b)
```

```{r supfig_mass2, fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
multipanels_mass2(lambda_vals2b)
```

```{r supfig_mass3, fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
multipanels_mass2(lambda_vals3b)
```

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_mass2(lambda_vals1b, ylim = c(0, ymax),
                  ylab = bquote(R[0]), hlines = 0:ymax, R0o = 5)
```

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_mass2(lambda_vals2b, ylim = c(0, ymax),
                  ylab = bquote(R[0]), hlines = 0:ymax, R0o = 5)
```

```{r fig.width = 8, fig.height = 7.875, margin1 = FALSE, margin3c = TRUE}
ymax <- 10
multipanels_mass2(lambda_vals3b, ylim = c(0, ymax),
                  ylab = bquote(R[0]), hlines = 0:ymax, R0o = 5)
```


#### Main text figure

```{r fig.width = 8, margin1 = FALSE, margin5 = TRUE}
mtext4 <- function(...) mtext(..., line = -.5, cex = 2, font = 2)
at_val <- 4.7
xmax <- 30

leg <- paste(seq(0, 100, 10), "%")
nb_mid <- 3

opar <- par(mfrow = 1:2)

plot_treatment_infected(attr(rho_val1, "params")$tau, rho_val1[[5]])
legend4("bottomleft", title = expression(paste("prevalence ", bar(X))),
        col = YlOrRd11, text.width = 0, legend = leg)

mtext4("a", at = - at_val / xmax)

plot_mass_treatment2(attr(lambda_vals1b, "params")$D_val / 24, lambda_vals1b[[5]])
legend4("bottomleft", title = "proportion treated", text.width = .5,
        col = c(YlOrRd11[1:3], rep("#FFFFFF", nb_mid), YlOrRd11[10:11]), 
        legend = c(leg[1:3], rep(".", nb_mid), leg[10:11]))

mtext4("b", at = at_val + xmax)

par(opar)
```

##### PDF output

Au format *Nature Medicine*:

```{r}
PDF <- TRUE

abline4pdf <- function(...) abline(..., col = "lightgrey")
abline3pdf <- function(...) abline4pdf(..., lwd = .5)

plot_treatment_infected_pdf <- function(x, y, x_axis = TRUE, y_axis = TRUE, ylim = 0:1,
                                        xlab = bquote("proportion" ~ tau ~ "of new cases that are treated"),
                                        ylab = ylab_rho,
                                        hlines1 = seq(0, 1, .1),
                                        hlines2 = seq(0, 1, .05),
                                        transform = 1) {
  plot(NA, xlim = 0:1, ylim = ylim, type = "n", xaxs = "i", yaxs = "i", axes = FALSE,
       ann = FALSE, xlab = xlab, ylab = ylab)
  ats <- seq(0, 1, .2)
  axis(1, ats, 100 * ats)
  axis(2, mgp = c(1, .75, 0))
  mtext(bquote("Proportion" ~ tau ~ "of new cases that are treated (%)"), 1, 1.5)
  mtext(ylab_rho, 2, 2, las = 0)
  abline3pdf(v = seq(0, 1, .05))
  abline4pdf(v = seq(0, 1, .1))
  abline4pdf(h = hlines1)
  abline3pdf(h = hlines2)
  walk2(y, YlOrRd11, ~ lines(x, .x * transform, col = .y))
  box(bty = "o")
}

legend4pdf <- function(...) {
  legend(lty = 1, lwd = 1, adj = c(1, .5), x.intersp = 4, ...)
}

plot_mass_treatment2_pdf <- function(x, y, x_axis = TRUE, y_axis = TRUE, ylim = 0:1,
                                xlab = "intake frequency (every D days)",
                                ylab = ylab_pi, hlines = seq(0, 1, .1), transform = 1,
                                xmin = 21, xmax = 56, xby = 7) {
  plot(rev(x), y, xlim = c(xmax, xmin), ylim = ylim, type = "n", ann = FALSE,
       xaxs = "i", yaxs = "i", axes = FALSE)
  
  axis(1, seq(xmin, xmax, xby))
  axis(2, mgp = c(1, .75, 0))
  mtext("Intake frequency (every D days)", 1, 1.5)
  mtext(ylab_pi, 2, 2, las = 0)
  
  abline(v = seq(xmin, xmax, xby), col = "grey")
  abline(h = hlines, col = "grey")
  walk2(seq(0, 1, .1), YlOrRd11,
        ~ lines(rev(x), rev((1 - (1 - min_of_following(y)) * .x) * transform),
                col = .y))
  box(bty = "o")
}

#######################################################################################

if (PDF) {
  pdf("figure6_new.pdf",
      width = 121 / 25.4, height = 60.5 / 25.4,
      family = "Helvetica", pointsize = 6)
}
  
mtext4pdf <- function(...) mtext(..., line = -.5, cex = 1.5, font = 2)
at_val <- 4.7
xmax <- 8 * 7
xmin <- 3 * 7

leg <- paste(seq(0, 100, 10), "%")
nb_mid <- 3

opar <- par(mfrow = 1:2, mgp = c(1.5, .5, 0), plt = c(.15, .95, .13, .93), las = 1)

plot_treatment_infected_pdf(attr(rho_val1, "params")$tau, rho_val1[[5]])
legend4pdf("bottomleft", title = expression(paste("Prevalence ", bar(X))), text.width = 0,
        col = YlOrRd11,
        legend = leg,
        bg = "white")

mtext4pdf("a", at = - at_val / (xmax - xmin))

#plot_mass_treatment_pdf(attr(lambda_vals, "params")$D_val / 24, lambda_vals[[5]])
plot_mass_treatment2_pdf(attr(lambda_vals1b, "params")$D_val / 24, lambda_vals1b[[5]])
legend4pdf("bottomleft", title = expression(paste("Proportion ", mu, " treated")), text.width = .5,
           col = YlOrRd11,
           legend = leg,
#        col = c(YlOrRd11[1:5], rep("#FFFFFF", nb_mid), YlOrRd11[10:11]), 
#        legend = c(leg[1:5], rep(".         .     ", nb_mid), leg[10:11]),
        bg = "white")

mtext4pdf("b", at = at_val + xmax)

par(opar)

if (PDF) dev.off()
```


#### TIFF output

```{r}
TIFF <- TRUE

if (TIFF) {
  tiff("figure6_900.tif",
       width = 121, height = 60.5, units = "mm", pointsize = 6, res = 900,
       family = "Helvetica")
}
  
mtext4pdf <- function(...) mtext(..., line = -.5, cex = 1.5, font = 2)
at_val <- 4.7
xmax <- 8 * 7
xmin <- 3 * 7

leg <- paste(seq(0, 100, 10), "%")
nb_mid <- 3

opar <- par(mfrow = 1:2, mgp = c(1.5, .5, 0), plt = c(.15, .95, .13, .93), las = 1)

plot_treatment_infected_pdf(attr(rho_val1, "params")$tau, rho_val1[[5]])
legend4pdf("bottomleft", title = expression(paste("Prevalence ", bar(X))), text.width = 0,
        col = YlOrRd11,
        legend = leg,
        bg = "white")

mtext4pdf("a", at = - at_val / (xmax - xmin))

plot_mass_treatment2_pdf(attr(lambda_vals1b, "params")$D_val / 24, lambda_vals1b[[5]])
legend4pdf("bottomleft", title = expression(paste("Proportion ", mu, " treated")), text.width = .5,
           col = YlOrRd11,
           legend = leg,
        bg = "white")

mtext4pdf("b", at = at_val + xmax)

par(opar)

if (TIFF) dev.off()
```

