---
title: "Mosquito-mediated epidemiological effects of treatment"
number-sections: true
format:
  html:
    toc: true
    toc-depth: 4
editor: source
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: the-american-naturalist.csl
---

```{r include = FALSE, eval = FALSE}
knitr::purl("new_model.qmd", "tmp.R", documentation = FALSE)
source("tmp.R")
file.remove("tmp.R")
```

```{r include = FALSE, purl = FALSE}
par2 <- function(...) par(..., mgp = c(1.5, .5, 0), bty = "n")

knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par2(plt = c(.105, .97, .15, .95)) else NULL
  })

eps <- .8
knitr::opts_chunk$set(margin1    = TRUE,
                      fig.retina = 2,
                      fig.align  = "center",
                      fig.height = eps * 5, # default is 5
                      fig.width = eps * 7) # default is 7
```


## Parameters

The path to the data:

```{r}
path2data <- paste0(Sys.getenv("HOME"), "/Library/CloudStorage/",
                    "OneDrive-OxfordUniversityClinicalResearchUnit/",
                    "GitHub/choisy/malarone/")
```

```{r include = FALSE}
make_path <- function(x) paste0(path2data, "cache/", x)
file_exists <- function(x) file.exists(make_path(x))
readRDS3 <- function(x) readRDS(make_path(x))
saveRDS2 <- function(object, file) saveRDS(object, make_path(file))
```

Whether or not to force the effects to be equal to $0$ at $t = 28$ days (in practice it
does not change things much anyway):

```{r}
t28_0 <- TRUE
```

Colors:

```{r}
color1 <- 2
color2 <- 3
color3 <- 4
```

Line width:

```{r}
lwd_val <- 2
```

Plot labels:

```{r}
xlab_pda <- "time post drug administration (hours)"
```


## Packages

```{r message = FALSE}
library(dplyr)
library(purrr)
```

## Functions

Tuning some base functions:

```{r}
seq2 <- function(...) seq(..., le = 512)
lines2 <- function(...) lines(..., lwd = lwd_val)
lines3 <- function(...) lines2(..., col = color1)
readRDS2 <- function(f, ...) tibble::as_tibble(readRDS(paste0(path2data, f), ...))
plot2 <- function(...) plot(..., col = color3, pch = 3, lwd = lwd_val)
```

A function that replace the `NA` values of a vector by `0`:

```{r}
NAto0 <- function(x) {
  x[is.na(x)] <- 0
  x
}
```




## Theoretical model

### Parasite load and infectiousness

#### 1 bite, all human treated

Call $t$ the **time post drug administration** in treated humans and $\tau$ the
**time post mosquito infection**. Call $\alpha(t, \tau)$ the
**transmission blocking activity** (TBA) defined as the percentage reduction of the
number of mosquitoes infected consecutively to 1 blood meal at times $t$ and $\tau$.
Call $\beta(t, \tau)$ the **transmission reducing activity** (TRA) defined as the
percentage reduction of the parasite load in infected mosquitoes consecutively to 1
blood meal at times $t$ and $\tau$. The mean reduction factor of the parasite load due
to TBA and TRA and across the population of infected mosquitoes at times $t$ and $\tau$
then reads

$$
(1 - \alpha(t, \tau))(1 - \beta(t, \tau)).
$$

Call $h(t)$ is the density of probability of the
**distribution of times post drug administration** in the population of treated humans.
The mean reduction factor $l(\tau)$ of parasite load acroos the population of infected
mosquitoes due to
**1 bite on treated people at time $\tau$** then reads

$$
l(\tau) = \int_0^{+\infty}(1 - \alpha(t, \tau))(1 - \beta(t, \tau))h(t)dt.
$$

#### Multiple bites, $p\%$ human treated

Call $a$ the **mosquito bite rate**, $p$ the
**proportion of the human population that receives treatment** and $\lambda(\tau)$ the
**mean cumulative reduction factor of the parasite load due to all the bites that have occurred until time $\tau$**.
The fact that each bite on a treated human reduces the parasite load of
$(1 - l(\tau)) \%$ is then conveyed by the following differential equation

$$
\frac{d\lambda}{d\tau} = -(1 - l(\tau)) a p \lambda(\tau).
$$

If we now assume that mosquitoes get infected by bites on untreated humans only then

$$
\lambda(0) = 1,
$$

which leads to the following expression of the mean cumulative reduction factor of the
parasite load at time $\tau$

$$
\lambda(\tau) = \exp\left[-ap\int_0^\tau(1 - l(u))du\right]
$$

#### Mean infectiousness reduction

Call $\mu(\lambda)$ the function that converts a **reduction factor of parasite load**
$\lambda$ into a **reduction factor of mosquito infectiousness** and $g$ the
**mosquito death rate** assumed to be independent of the mosquito's age. The mean
reduction factor of mosquito infectiousness in the total mosquito population due to TBA
and TRA can then be expressed as

$$
\theta = \int_0^{+\infty}e^{-g\tau}\mu(\lambda(\tau))d\tau.
$$

### Parasite maturation and EIP

The EIP is the time at which the parasite reaches maturity for transmission. Call
$z(\tau)$ the dimensionless **level of parasite maturity** at time $\tau$ post
infection and assume it increases from some arbitrary value 0 right after infection:

$$
z(0) = 0
$$

to some arbitrary value 1 once maturity for transmission is reached:

$$
z(\mbox{EIP}) = 1
$$

at a **maturation rate** $\nu$ that is independent of the level of maturity (but that
can depend of the time $\tau$ post infection due in particular to bites on treated
humans):

$$
\frac{dz}{d\tau} = \nu(\tau).
$$

Note that the values 0 and 1 are just convenient choices here and that their exact
values do not matter for the analysis.

#### 1 bite, all human treated

Call $\gamma(t, \tau)$ the **reduction factor of the maturation rate** consecutively to
1 blood meal at times $t$ and $\tau$. The mean reduction factor $q(\tau)$ of the
maturation rate across the population of infected mosquitoes due to
**1 bite on treated people at time $\tau$** then reads

$$
q(\tau) = \int_0^{+\infty} \gamma(t, \tau) h(t) dt
$$

#### Multiple bites, $p\%$ human treated

In a similar way as done above for the parasite load, we can express the fact that each
bite on treated human reduces the maturation rate by $1 - q(\tau)$ by the following
differential equation

$$
\frac{d\nu}{d\tau} = -(1 - q(\tau))ap\nu(\tau)
$$
Under the same assumption as above that mosquitoes get infected by bites on untreated
humans only then

$$
\nu(0) = \frac{1}{n}
$$

leading to

$$
\nu(\tau) = \frac{1}{n} \exp\left[-ap\int_0^{\tau}(1 - q(u))du\right]
$$

where $n$ is the value of the **EIP in absence of treatment** in the human population.
Now that we have this, we can implicitly express the average EIP in presence of the
treatment in the human population as the value $T$ satisfying 

$$
\int_0^{T}\nu(\tau)d\tau = 1
$$

### $R_0$ reduction effect

Putting previous results together, if $R_0$ is the value of the basic reproduction
number in absence of treatment, then, in the presence of treatment in the population,
its expression becomes

$$
\theta e^{-(T - n)g} R_0
$$

## Experimental data

Let's use the following 4-parameter functional relationship for $\alpha(t, \tau)$,
$\beta(t, \tau)$ and $\gamma(t, \tau)$:

$$
f(t,\tau) = A \cdot (1 - e^{-kt})\cdot e^{-(dt + r\tau)}
$$

whereby, as function of $t$, the drug effects are characterized by an initial
saturating increase (captured by $k$) followed by an exponential decrease (captured by
$d$) reflecting the PK/PD of the drug and, as function of $\tau$, the effect is
characterized by an exponential decrease (captured by $r$) simply reflecting the growth
of the parasite (and thus the decrease of the relative concentration of the drug in
each blood meal). Parameter $A$ account for the maximum value reached.

```{r}
f <- function(t, tau, A, k, d, r) A * (1 - exp(-k * t)) * exp(-(d * t + r * tau))
```

A version of the model at 4 dpi:

```{r}
f4dpi <- function(t, A, k, d) A * (1 - exp(-k * t)) * exp(-d * t)
```

A function that looks for initial values at 4 dpi:

```{r}
initial_values <- function(x, y) {
  df <- tibble(x, y) |> 
    arrange(x)

  df2 <- df |> 
    group_by(x) |> 
    summarise(y = mean(y))
  
  A <- df2 |> 
    pull(y) |> 
    max()
  
  xs <- df2 |> 
    filter(y == A) |> 
    pull(x)

  k <- df |> 
    filter(x < first(xs)) |> 
    lm(y ~ -1 + x, data = _) |> 
    coef() |> 
    last()
  
  d <- df |>
    filter(x > last(xs)) |> 
    filter(y > 0) |> 
    mutate(across(y, log)) |> 
    lm(y ~ x, data = _) |> 
    coef() |> 
    last()
  
  c(A = A, k = k / A, d = -d)
}
```

A function that calibrates the model at 4 dpi with least sum of squares:

```{r}
calibrate4dpi <- function(x, y) {
  optim(initial_values(x, y), function(p) sum((y - f4dpi(x, p[1], p[2], p[3]))^2))$par
}
```


### Mosquitoes with oocysts

Below, we generate the function $\alpha(t)$ from the experimental data of Figure 2a.
First we need a function that reads experimental data for the treatment group:

```{r}
make_reading_exp_data <- function() {
  hours <- c(0, 5, 48, 72)
  days <- c(5, 7, 14, 28)
  hash <- c(setNames(hours, paste0("H", hours)),
            setNames(24 * days, paste0("D", days)))
  function(file) {
    file |>
      readRDS2() |> 
      filter(treatment == "AL+AP") |> 
      mutate(time_hours = hash[stringr::str_remove(time_points, "0")],
             across(Inhibition, ~ .x / 100))
  }
}

reading_exp_data <- make_reading_exp_data()
```

```{r}
alpha_data <- "figure2a.rds" |>
  reading_exp_data() |> 
  rename(t = time_hours, alpha = Inhibition) |> 
  select(t, alpha)
```

```{r}
x <- seq2(0, 700)
parms <- with(alpha_data, calibrate4dpi(t, alpha))
with(alpha_data, plot2(jitter(t), alpha, xlab = xlab_pda,
                       ylab = bquote("reduction" ~ alpha ~
                                       "of mosquitoes with oocysts")))

lines3(x, with(as.list(parms), f4dpi(x, A, k, d)))
abline(h = 0)
```

Stratisfied bootstrapping:

```{r}
stratified_bootstrap <- function(df, N = 1000) {
  map(1:N, ~ df |> 
               group_by(t) |> 
               group_split() |> 
               map(slice_sample, prop = 1, replace = TRUE) |> 
               bind_rows())
}
```

```{r}
alpha_model <- alpha_data |> 
  stratified_bootstrap(3) |> 
  map(~ with(.x, calibrate4dpi(t, alpha)))
```

```{r}
alpha_model |> 
  map_dfc(~ with(as.list(.x), f4dpi(x, A, k, d))) |> 
  walk(~ lines(x, .x, col = 2))
```

