---
title: "Mosquito-mediated epidemiological effects of treatment"
number-sections: true
format:
  html:
    toc: true
    toc-depth: 4
editor: source
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: the-american-naturalist.csl
---

```{r include = FALSE, eval = FALSE}
# The code that adds the .nojekyll file to the GitHub repository:
file <- ".nojekyll"
file.create(file)
gert::git_add(file)
gert::git_commit("Adding the .nojekyll file")
gert::git_push()
rm(file)

knitr::purl("new_model.qmd", "tmp.R", documentation = FALSE)
source("tmp.R")
file.remove("tmp.R")
```

```{r include = FALSE, eval = FALSE}
# To run all the code except the cached chunks:
run_all <- function() {
  knitr::purl("new_model.qmd", "tmp.R", documentation = FALSE)
  source("tmp.R")
  file.remove("tmp.R")  
}

rerun_all <- function() {
  rm(list  = grep("run_all", ls(envir = .GlobalEnv), value = TRUE, invert = TRUE),
     envir = .GlobalEnv)
  run_all()
}
```


## Parameters

The path to the data:

```{r}
if (Sys.getenv("USER") == "MarcChoisy" | Sys.getenv("USER") == "marc") {
  path2data <- paste0(Sys.getenv("HOME"), "/Library/CloudStorage/",
                      "OneDrive-OxfordUniversityClinicalResearchUnit/",
                      "GitHub/choisy/malarone/")
}
```

```{r include  = FALSE}
path2cache <- paste0(path2data, "cache/")
if (! dir.exists(path2cache)) dir.create(path2cache)

make_path <- function(x) paste0(path2cache, x)
file_exists <- function(x) file.exists(make_path(x))
readRDScache <- function(x) readRDS(make_path(x))
saveRDScache <- function(object, file) saveRDS(object, make_path(file))
remove_cache <- function(x) {
  file.remove(grep(x, dir(path2cache, full.names = TRUE), value = TRUE))
}
```

```{r include = FALSE, purl = FALSE}
par2 <- function(...) par(..., mgp = c(1.5, .5, 0), bty = "n")

knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par2(plt = c(.105, .97, .15, .95)) else NULL
  },
  margin2 = function(before, options, envir) {
    if (before) par2(plt = c(.05, .92, .15, .95)) else NULL
  })

eps <- .8
knitr::opts_chunk$set(margin1    = TRUE,
                      fig.retina = 2,
                      fig.align  = "center",
                      fig.height = eps * 5, # default is 5
                      fig.width = eps * 7) # default is 7
```

Line width:

```{r}
lwd_val <- 2
```


## Packages

```{r message = FALSE}
library(dplyr)
library(purrr)
```


## Functions

Tuning `polygon()`:

```{r}
polygon2 <- function(x, y1, y2, ...) {
  polygon(c(x, rev(x)), c(y1, rev(y2)), border = NA, ...)
}
```

A function to read the experimental data:

```{r}
reading_exp_data <- function(file) {
  hours <- c(0, 5, 48, 72)
  days <- c(5, 7, 14, 28)
  hash <- c(setNames(hours, paste0("H", hours)),
            setNames(24 * days, paste0("D", days)))
  
  path2data |> 
    paste0(file) |> 
    readRDS() |> 
    tibble::as_tibble() |> 
    filter(treatment == "AL+AP") |> 
    mutate(t = hash[stringr::str_remove(time_points, "0")],
           value = Inhibition / 100) |> 
    select(t, value)
}
```

Stratified bootstrapping:

```{r}
stratified_bootstrap <- function(df, N) {
  map(1:ceiling(N), ~ df |> 
                        group_by(t) |> 
                        group_split() |> 
                        map(slice_sample, prop = 1, replace = TRUE) |> 
                        bind_rows())
}
```


## Theoretical model

### Parasite load and infectiousness

#### 1 bite, all human treated

Call $t$ the **time post drug administration** in treated humans and $\tau$ the
**time post mosquito infection**. Call $\alpha(t, \tau)$ the
**transmission blocking activity** (TBA) defined as the percentage reduction of the
number of mosquitoes infected consecutively to 1 blood meal at times $t$ and $\tau$.
Call $\beta(t, \tau)$ the **transmission reducing activity** (TRA) defined as the
percentage reduction of the parasite load in infected mosquitoes consecutively to 1
blood meal at times $t$ and $\tau$. The mean reduction factor of the parasite load due
to TBA and TRA and across the population of infected mosquitoes at times $t$ and $\tau$
then reads

$$
(1 - \alpha(t, \tau))(1 - \beta(t, \tau)).
$$

Call $h(t)$ is the density of probability of the
**distribution of times post drug administration** in the population of treated humans.
The mean reduction factor $l(\tau)$ of parasite load acroos the population of infected
mosquitoes due to
**1 bite on treated people at time $\tau$** then reads

$$
l(\tau) = \int_0^{+\infty}(1 - \alpha(t, \tau))(1 - \beta(t, \tau))h(t)dt.
$$

#### Multiple bites, $p\%$ human treated

Call $a$ the **mosquito bite rate**, $p$ the
**proportion of the human population that receives treatment** and $\lambda(\tau)$ the
**mean cumulative reduction factor of the parasite load due to all the bites that have occurred until time $\tau$**.
The fact that each bite on a treated human reduces the parasite load of
$(1 - l(\tau)) \%$ is then conveyed by the following differential equation

$$
\frac{d\lambda}{d\tau} = -(1 - l(\tau)) a p \lambda(\tau).
$$

If we now assume that mosquitoes get infected by bites on untreated humans only then

$$
\lambda(0) = 1,
$$

which leads to the following expression of the mean cumulative reduction factor of the
parasite load at time $\tau$

$$
\lambda(\tau) = \exp\left[-ap\int_0^\tau(1 - l(u))du\right]
$$

#### Mean infectiousness reduction

Call $\zeta(\lambda)$ the function that converts a **reduction factor of parasite load**
$\lambda$ into a **reduction factor of mosquito infectiousness** and $\mu$ the
**mosquito death rate** assumed to be independent of the mosquito's age. The mean
reduction factor of mosquito infectiousness in the total mosquito population due to TBA
and TRA can then be expressed as

$$
\theta = \int_0^{+\infty}e^{-\mu\tau}\zeta(\lambda(\tau))d\tau.
$$

### Parasite maturation and EIP

The EIP is the time at which the parasite reaches maturity for transmission. Call
$z(\tau)$ the dimensionless **level of parasite maturity** at time $\tau$ post
infection and assume it increases from some arbitrary value 0 right after infection:

$$
z(0) = 0
$$

to some arbitrary value 1 once maturity for transmission is reached:

$$
z(\mbox{EIP}) = 1
$$

at a **maturation rate** $\nu$ that is independent of the level of maturity (but that
can depend of the time $\tau$ post infection due in particular to bites on treated
humans):

$$
\frac{dz}{d\tau} = \nu(\tau).
$$

Note that the values 0 and 1 are just convenient choices here and that their exact
values do not matter for the analysis.

#### 1 bite, all human treated

Call $\gamma(t, \tau)$ the **reduction factor of the maturation rate** consecutively to
1 blood meal at times $t$ and $\tau$. The mean reduction factor $q(\tau)$ of the
maturation rate across the population of infected mosquitoes due to
**1 bite on treated people at time $\tau$** then reads

$$
q(\tau) = \int_0^{+\infty} \gamma(t, \tau) h(t) dt
$$

#### Multiple bites, $p\%$ human treated

In a similar way as done above for the parasite load, we can express the fact that each
bite on treated human reduces the maturation rate by $1 - q(\tau)$ by the following
differential equation

$$
\frac{d\nu}{d\tau} = -(1 - q(\tau))ap\nu(\tau)
$$
Under the same assumption as above that mosquitoes get infected by bites on untreated
humans only then

$$
\nu(0) = \frac{1}{n}
$$

leading to

$$
\nu(\tau) = \frac{1}{n} \exp\left[-ap\int_0^{\tau}(1 - q(u))du\right]
$$

where $n$ is the value of the **EIP in absence of treatment** in the human population.
Now that we have this, we can implicitly express the average EIP in presence of the
treatment in the human population as the value $T$ satisfying 

$$
\int_0^{T}\nu(\tau)d\tau = 1
$$

### $R_0$ reduction effect

Putting previous results together, if $R_0$ is the value of the basic reproduction
number in absence of treatment, then, in the presence of treatment in the population,
its expression becomes

$$
\theta e^{-(T - n)\mu} R_0
$$


## Experimental data

Let's use the following 4-parameter functional relationship for $\alpha(t, \tau)$,
$\beta(t, \tau)$ and $\gamma(t, \tau)$:

$$
f(t,\tau) = A \cdot (1 - e^{-kt})\cdot e^{-(dt + r\tau)}
$$

whereby, as function of $t$, the drug effects are characterized by an initial
saturating increase (captured by $k$) followed by an exponential decrease (captured by
$d$) reflecting the PK/PD of the drug and, as function of $\tau$, the effect is
characterized by an exponential decrease (captured by $r$) simply reflecting the growth
of the parasite (and thus the decrease of the relative concentration of the drug in
each blood meal). Parameter $A$ account for the maximum value reached. Given that the
efficacy of a drug depends on the concentration of the drug respective to the size of
the parasite, one can argue $r = d$, which simplifies the formula:

$$
f(t,\tau) = A \cdot (1 - e^{-kt})\cdot e^{-d\cdot(t + \tau)}
$$

A version of the model at 4 dpi:

```{r}
f4dpi <- function(t, A, k, d) A * (1 - exp(-k * t)) * exp(-d * t)
```

A function that looks for initial values at 4 dpi:

```{r}
initial_values <- function(x, y) {
  df <- tibble(x, y) |> 
    arrange(x)

  df2 <- df |> 
    group_by(x) |> 
    summarise(y = mean(y))
  
  A <- df2 |> 
    pull(y) |> 
    max()
  
  xs <- df2 |> 
    filter(y == A) |> 
    pull(x)

  k <- df |> 
    filter(x < first(xs)) |> 
    lm(y ~ -1 + x, data = _) |> 
    coef() |> 
    last()
  
  d <- df |>
    filter(x > last(xs)) |> 
    filter(y > 0) |> 
    mutate(across(y, log)) |> 
    lm(y ~ x, data = _) |> 
    coef() |> 
    last()
  
  c(A = A, k = k / A, d = -d)
}
```

A function that calibrates the model at 4 dpi with least sum of squares:

```{r}
calibrate4dpi <- function(x, y) {
  optim(initial_values(x, y), function(p) sum((y - f4dpi(x, p[1], p[2], p[3]))^2))$par
}
```


### Mosquitoes with oocysts

Below, we generate the function $\alpha(t)$ from the experimental data of Figure 2a.
First we need a function that reads experimental data for the treatment group:

```{r}
alpha_data <- reading_exp_data("figure2a.rds")
```

Fitting the model to the data:

```{r}
alpha_fit <- with(alpha_data, calibrate4dpi(t, value))
```

```{r}
make_effect <- function(A, k, d) {
  A <- A * exp(4 * d)
  function(t, tau) A * (1 - exp(-k * t)) * exp(-d * (t + tau))
}
```

```{r}
alpha <- with(as.list(alpha_fit), make_effect(A, k, d))
```

```{r fig.width = 7}
t <- seq(0, 350, le = 100)
tau <- seq(0, 350, le = 10)
alpha3d <- tidyr::expand_grid(t, tau) |> 
  mutate(value = alpha(t, tau)) |> 
  pull(value) |>
  matrix(10) |> 
  persp(tau, t, z = _, theta = 45, zlab = "TBA alpha")
```

```{r}
x_val <- seq(0, 700)
plot(x_val, alpha(x_val, tau = 0), type = "n",
     xlab = "time post drug administration (hours)",
     ylab = bquote("reduction" ~ alpha ~ "of mosquitoes with oocysts"))
walk(seq(0, 700, 4), ~ lines(x_val, alpha(x_val, tau = .x)))
lines(x_val, alpha(x_val, tau = 4), col = 2, lwd = lwd_val)
```

It takes (1'20" on the laptop):

```{r}
calibration <- function(data, N = 1e4) {
  model <- data |> 
    stratified_bootstrap(1.1 * N) |> 
    map(~ with(.x, safely(calibrate4dpi)(t, value)))

  model[map_lgl(model, ~ is.null(.x$error))] |>
    map("result") |> 
    sample(N)
}
```

```{r eval = FALSE, purl = FALSE}
alpha_model <- calibration(alpha_data)
```

```{r include = FALSE}
if (file_exists("alpha_model.rds")) {
  alpha_model <- readRDScache("alpha_model.rds")
} else {
  alpha_model <- calibration(alpha_data)
  saveRDScache(alpha_model, "alpha_model.rds")
}
```

```{r}
x_max <- 700
x <- seq(0, 1.05 * x_max)

simulate <- function(model, xs = x) {
  model |> 
    set_names() |> 
    map_dfc(~ with(as.list(.x), f4dpi(xs, A, k, d)))
}
```

```{r}
alpha_simulations <- simulate(alpha_model)
```

```{r}
plot_data_model <- function(data, simulations, ylabel, col_data = 4, col_model = 2) {
  with(data, plot(t, value, type = "n",
                  xlim = c(0, x_max), ylim = c(min(data$value), 1),
                  xlab = "time post drug administration (hours)",
                  ylab = ylabel))
  abline(h = 0)
  
  simulations |> 
    t() |> 
    as.data.frame() |> 
    map_dfr(quantile, c(.025, .5, .975)) |> 
    with({
      polygon2(x, `2.5%`, `97.5%`, col = adjustcolor(col_model, .1))
      lines(x, `50%`, lwd = lwd_val, col = col_model)
    })
  
  with(data, points(jitter(t), value, col = col_data, lwd = lwd_val, pch = 3))
}
```

```{r}
plot_data_model(alpha_data, alpha_simulations,
                bquote("reduction" ~ alpha ~ "of mosquitoes with oocysts"))

lines(x, with(as.list(alpha_fit), f4dpi(x, A, k, d)))
```


### Number of oocysts

```{r}
beta_data <- reading_exp_data("figure2c.rds")
```

Fitting the model to the data:

```{r}
beta_fit <- with(beta_data, calibrate4dpi(t, value))
```

```{r eval = FALSE, purl = FALSE}
beta_model <- calibration(beta_data)
```

```{r include = FALSE}
if (file_exists("beta_model.rds")) {
  beta_model <- readRDScache("beta_model.rds")
} else {
  beta_model <- calibration(beta_data)
  saveRDScache(beta_model, "beta_model.rds")
}
```

```{r}
beta_simulations <- simulate(beta_model)
```

```{r}
plot_data_model(beta_data, beta_simulations,
                bquote("reduction" ~ beta ~ "of the number of oocysts"))

lines(x, with(as.list(beta_fit), f4dpi(x, A, k, d)))
```


### Converting $\alpha$ and $\beta$ into reduction of infectiousness

#### Old data

Here we are computing the function $\mu$ that converts a reduction in oocyst load into
a reduction in infectiousness. 

```{r message = FALSE}
oocysts_control <- path2data |>
  paste0("ASAAP4b_oocyst2.txt") |>
  readr::read_table() |> 
  filter(day_of_dissection == "7dpi",
         time_points %in% c("Control", "Unfed"),
         total_oocysts > 0) |> 
  select(total_oocysts) |> 
  na.exclude()
```

```{r}
total_oocysts <- oocysts_control$total_oocysts
  
hist(total_oocysts, main = NA,
     xlab = "oocyst load L in the controls", ylab = "frequency")
```

```{r}
max_load <- max(total_oocysts)

hist(total_oocysts, 0:max_load + .5, main = NA,
     xlab = "oocyst load L in the controls", ylab = "frequency")
```

```{r}
oocysts_control |> 
  group_by(total_oocysts) |> 
  tally() |> 
  mutate(cum_n = cumsum(n) / sum(n)) |> 
  with(plot(total_oocysts, cum_n, type = "s", ylim = 0:1,
            xlab = "oocyst load L", ylab = "cumulative proportion of samples"))

p_par <- 1 / mean(total_oocysts)
x_s <- 1:max_load
lines(x_s, cumsum(dgeom(x_s, p_par)), type = "s", col = 2)
```

#### New data

```{r message = FALSE}
bompard2020 <- path2data |>
  paste0("doi_10_5061_dryad_9p8cz8wfh__v20210123/act2datasetRpcr.txt") |>
  readr::read_table()
```

```{r}
bompard2020 |> 
  filter(Total > 0) |> 
  pull(Total) |> 
  hist(0:800, xlab = "oocyst load L", ylab = "frequency", main = NA)
```

```{r}
bompard2020 |> 
  filter(0 < Total, Total < 50) |> 
  pull(Total) |> 
  hist(0:50, xlab = "oocyst load L", ylab = "frequency", main = NA)
```

```{r}
negbinom <- bompard2020 |> 
  pull(Total) |> 
  fitdistr("negative binomial")
```

```{r}
fit_zi_nbinom <- function(x) {
  require(glmmTMB)
  fit <- glmmTMB(y ~ 1, data.frame(y = x), nbinom2, ~ 1)
  c(pi   = plogis(fixef(fit)$zi[[1]]),
    size = sigma(fit),
    mu   = predict(fit, type = "conditional")[1])
}

fit_zi_nbinom(c(rep(0, 2000), rnbinom(8000, size = 28.75, mu = 1.8)))
```

```{r}
bompard2020 |> 
  pull(Total) |> 
  fit_zi_nbinom()
```

```{r}
bompard2020 |> 
  group_by(Total) |> 
  tally() |> 
  mutate(cum_n = cumsum(n) / sum(n)) |> 
  with(plot(Total, cum_n, type = "s", xlim = c(0, 50), ylim = c(.85, 1),
            xlab = "oocyst load L", ylab = "cumulative proportion of samples"))
```




```{r}
bompard2020 |> 
  group_by(Total) |> 
  tally() |> 
  mutate(cum_n = cumsum(n) / sum(n)) |> 
  with(plot(Total, cum_n, type = "s", ylim = 0:1,
            xlab = "oocyst load L", ylab = "cumulative proportion of samples"))

p_par2 <- 1 / mean(bompard2020$Total)
x_s <- 1:max_load
lines(x_s, cumsum(dgeom(x_s, p_par2)), type = "s", col = 2)
lines(1:800, cumsum(dnbinom(1:800, 1.5, mu = 10.95123598)), col = 4)
lines(1:800, cumsum(dnbinom(1:800, size = 1 / sigma(fit), mu = predict(fit, type = "conditional")[1])))
```



Figure 2B from @aleshnick_experimental_2020 suggests a non-linear relationship $h$
between the sporozoite load and the transmission efficiency $b$ of *P. yoelii* from
*An. stephensi* mosquito to mice with 10,000 salivary gland sporozoites being
significantly more likely to initiate a malaria infection. Furthermore, Figure 1 from
@miura_strong_2019 documents a strong linear relationship between the numbers of
*P. falciparum* oocysts at 8 dpi and sporozoites at 16 dpi in *An. stephensi*
mosquitoes with about 1,000 sporozoites per oocyst (966, 95\% CI: 893–1038)).

Let's load the data from Figure 2B of @aleshnick_experimental_2020 on the relationship
between sporozoite load and the proportion of successful infections, converting it
into a relationship with oocyst load using @miura_strong_2019's 1000 factor.

```{r}
aleshnick2020 <- path2data |> 
  paste0("ppat.1008181.s006.xlsx") |> 
  readxl::read_excel() |> 
  mutate(load = `Salivary Gland Load` / 1000) |> 
  rename(infection = `Malaria Infection`) |> 
  select(load, infection)
```

A 2-parameter exponential saturating function of the probability of transmission as a
function of the parasite load:

$$
g(L) = B\cdot (1 - e^{-sL})
$$

```{r}
g <- function(L, B, s) B * (1 - exp(-s * L))
```

```{r}
th <- 25

B <- aleshnick2020 |> 
  filter(load > th) |> 
  pull(infection) |> 
  mean()

s <- aleshnick2020 |> 
  filter(load < th) |> 
  arrange(load) |> 
  map(~ zoo::rollmean(.x, 10)) |> 
  as_tibble() |> 
  lm(infection ~ -1 + load, data = _) |> 
  coef()

mLL <- function(x) {
  -sum(with(aleshnick2020, dbinom(infection, 1, g(load, exp(x[1]), exp(x[2])), TRUE)))
}

params <- exp(optim(c(B = log(B), s = log(s)), mLL)$par)
g_calibrated <- function(x) g(x, params[1], params[2])
```

```{r}
load_vals <- seq(0, max_load, le = 512)

plot(load_vals, g_calibrated(load_vals), type = "l", lwd = lwd_val, col = 2,
     xlim = c(0, max_load), ylim = c(0, .5),
     xlab = "oocyst load L", ylab = "probability of transmission")
```

$$
\zeta(\lambda) = \sum_{L = 1}^{+\infty}\frac{g(\lambda L)}{g(L)}\omega(L)
$$

$$
\zeta(\lambda) = \sum_{L = 1}^{+\infty}\frac{1 - e^{-\lambda L}}{1 - e^{-L}}(1-p_o)^{L-1}p_o
$$

Approximation:

$$
\zeta(\lambda) \simeq \zeta_w(\lambda) = \frac{\lambda}{w + (1 - w)\lambda}
$$

with a constant $w$.

```{r}
mu1 <- function(lambda) {
  sum(map_dbl(1:1000,
              ~ dgeom(.x - 1, p_par) * g_calibrated(lambda * .x) / g_calibrated(.x)))
}

mu <- function(lambda) map_dbl(lambda, mu1)
```

```{r eval = FALSE, purl = FALSE}
lambda_vals <- seq(0, 1, le = 512)
mu_vals <- mu(lambda_vals)
```

```{r include = FALSE}
lambda_vals <- seq(0, 1, le = 512)
if (file_exists("mu_vals.rds")) {
  mu_vals <- readRDScache("mu_vals.rds")
} else {
  mu_vals <- mu(lambda_vals)
  saveRDScache(mu_vals, "mu_vals.rds")
}
```

```{r fig.height = eps * 6, fig.width = eps * 8, margin1 = FALSE, margin2 = TRUE}
#| label: fig-b_reduction
#| fig-cap: the conversion of a reduction on the sporozoite load $L$ into a reduction of the mosquito-to-human probability of transmission $b$ and $R_0$.

opar <- par(pty = "s", xpd = TRUE)
plot(lambda_vals, mu_vals, type = "l", lwd = lwd_val, col = 4, xaxs = "i", yaxs = "i",
     xlab = bquote("reducing factor of oocyst load"),
     ylab = "reducing factor of transmission")
segments(0, 0, 1, 1)
box(bty = "o")
w <- .36
lines(lambda_vals, lambda_vals / (w + (1 - w) * lambda_vals), col = 2)
par(opar)
```

Note that although the $\zeta$ function is not perfectly symmetric respective to the
second diagonal as the approximation is, the approximation with $w = 0.36$ does a
rather good job.

```{r}
l_4dpi <- function(alpha, beta, h) {
  integrate(function(t) (1 - alpha(t)) * (1 - beta(t)) * h(t), 0, Inf)
}
```

```{r}
lambda <- function(tau, a, p) {
  exp(-a * p * integrate(function(u) (1 - l_4dpi(u)), 0, tau))
}
```

```{r}
theta <- function(g) {
  integrate(function(tau) exp(-g * tau), 0, Inf)
}
```


### Size of oocysts at 7 dpi

```{r}
delta_data <- reading_exp_data("percentage_inhibition_size2.rds")
```

```{r eval = FALSE, purl = FALSE}
delta_model <- calibration(delta_data)
```

```{r include = FALSE}
if (file_exists("delta_model.rds")) {
  delta_model <- readRDScache("delta_model.rds")
} else {
  delta_model <- calibration(delta_data)
  saveRDScache(delta_model, "delta_model.rds")
}
```

```{r}
delta_simulations <- simulate(delta_model)
```

```{r}
plot_data_model(delta_data, delta_simulations,
                bquote("reduction" ~ delta ~ "of the size of oocysts"))
```

$$
\gamma(\delta) = (1 - \delta) + \frac{C}{D}\delta
$$

where $C$ is the size at 4 dpi and $D$ is the size increase from 4 to 7 dpi in the
controls.

```{r}
volume <- function(V0, Vmax, r, t) {
  Vmax * V0 * exp(r * t) / (Vmax - V0 + V0 * exp(r * t))
}

area <- function(v) (9 * pi / 16)^(1 / 3) * v^(2 / 3)
curve(volume(V0 = .1, Vmax = 10, r = .1, x), 0, 100)
```

